---
title: "ChatGPT Analysis of Mean Discharge in a River"
author: "Matt Ross"
date: "2023-01-23"
output:
  word_document: default
  html_document: default
---

# ChatGPT, the analyst

For this assignment you will be asking ChatGPT to conduct an analysis of mean river
discharge for any river station in the USA. You can use the Poudre River in 
Fort Collins if you want, but I encourage you to find other USGS stations that 
you are interested in. This assignment will have three prompts for ChatGPT, and
you will need to troubleshoot this code for each prompt. The prompts are:

1. *Using R, Download daily discharge data for XX river in YY city from 1990-2022 (you pick the river and 
the city).* You can choose to provide ChatGPT with an actual USGS station ID which 
will produce less errors, or you can see how well it figures out station ids 
(its bad at this for some reason). You may also need to provide it with specific
package recommendations (as I do in the videos).

2. How would I take the summertime (June 21 - Sept 21) annual mean of this data and plot it
with year on the x-axis and summertime mean flow on the y axis.

3. How would I Use a linear model to determine if summer annual mean discharge is
changing? 

4 (Bonus). Why might a linear model be a poor choice for this analysis? 


## Data Download

## Install the 'tidyverse' package
install.packages("tidyverse")

```{r, error = TRUE}
# Download the 'tidyverse' package

## ChatGPT asked me here to download tidyverse. I already have tidyverse downloaded. Not breaking R.

# Load the 'tidyverse' package
library(tidyverse)

# Download the data
data <- read.csv("https://waterdata.usgs.gov/nwis/dv?cb_00060=on&format=rdb&site_no=06730200&referred_module=sw&period=&begin_date=1990-01-01&end_date=2022-12-31")

# Clean up the data
data_cleaned <- data %>% 
  filter(!is.na(data$discharge)) %>% 
  mutate(date = as.Date(data$datetime, format = "%Y-%m-%d")) %>% 
  select(date, discharge)

# Write the data to a CSV file
write.csv(data_cleaned, file = "CacheLaPoudreRiver_Discharge_1990-2022.csv")

# This code will download the data from the USGS website, clean up the data by removing missing values and converting the date column to the Date class, and then write the cleaned data to a CSV file.
```

Fixing this. ChatGPT downloaded an absolutely awful ".csv" file from USGS for Boulder Creek. I decided to just download the thing myself using the dataRetrieval package.

```{r}
# Load the 'tidyverse' package
library(tidyverse)

# Load the dataRetrieval package
library(dataRetrieval)

# PoudreRiverData
PoudreRiverData <- readNWISdv("06752260", "00060", "1990-01-01", "2023-01-01")
```

```{r}
# Cleaning up the data
PoudreRiverData <- PoudreRiverData %>%
  mutate('Streamflow' = X_00060_00003) %>%
  filter(!is.na(Streamflow))
```

## Data analysis and visualization

```{r, error = TRUE}
library(tidyverse)

# Load the data
data <- read.csv("Cache_La_Poudre_River_Fort_Collins_Colorado_discharge.csv")

# Filter the data for summer months (June 21 - Sept 21)
summer_data <- data %>%
  filter(month(as.Date(date)) >= 6 & month(as.Date(date)) <= 9 & day(as.Date(date)) >= 21)

# Group the data by year
summer_data_by_year <- summer_data %>%
  group_by(year = year(as.Date(date))) %>%
  summarize(mean_flow = mean(discharge))

# Plot the data
ggplot(summer_data_by_year, aes(x = year, y = mean_flow)) +
  geom_line() +
  ggtitle("Summertime Annual Mean Flow of Cache La Poudre River in Fort Collins, Colorado") +
  xlab("Year") +
  ylab("Summertime Annual Mean Flow (cfs)")

# Note: In this code, the tidyverse library is used to perform the data filtering, grouping, and summarizing operations. The ggplot function is used to create the plot, with the x and y aesthetics specifying the year and summertime mean flow variables, respectively. The geom_line function is used to add a line to the plot, and the ggtitle, xlab, and ylab functions are used to add the title, x-axis label, and y-axis label to the plot, respectively.
```

Doing this in multiple code chunks just to have redundancy. Might clean this code up later, hehe.

```{r}
# ChatGPT did not use the 'separate' function properly.
PoudreRiverDataCleaned <- PoudreRiverData %>%
  separate(col = Date, into = c("year","month", "day"), sep = "-")
```

```{r}
PoudreRiverMonthDay <- PoudreRiverDataCleaned %>%
    unite(col = MonthDay, month, day, sep = "")
```

```{r}
PoudreRiverMonthDay$MonthDay <- as.numeric(PoudreRiverMonthDay$MonthDay)
```

```{r}
PoudreRiverSummerData <- PoudreRiverMonthDay %>%
  filter(MonthDay >= 621, MonthDay <= 921)
```

```{r}
Summerized <- PoudreRiverSummerData %>%
  group_by(year) %>%
  summarise(SummerAvg = mean(Streamflow))
```

```{r}
Summerized$year <- as.Date(Summerized$year, format = "%Y")
```

```{r}
ggplot(Summerized, aes(x = year, y = SummerAvg)) +
  geom_smooth(color = "lightblue", fill = "lightblue") +
  geom_point(color = "darkblue") +
  labs(title = "Average Summer Streamflow in the Cache la Poudre River, 1990-2022", subtitle = "USGS Station 6752260, Fort Collins, Colorado", x = "Year", y = "Streamflow (cfs)") +
  theme_light()
```

## Model output

To determine if summer annual mean discharge is changing over time, you can fit a linear regression model to the summertime mean flow data. You would need to install and load the necessary packages in R, such as tidyverse or lm() function from base R. Then, you would create a linear model with the summertime mean flow as the response variable and year as the predictor. Finally, you can use the summary() function to get a summary of the model, including the p-value of the t-test for the slope coefficient. If the p-value is less than a significance level (e.g. 0.05), then you can conclude that there is evidence that the summertime mean discharge is changing over time.

## Very well explained.

## Bonus - why linear models are dangerous time-series analysis tools. 

A linear model assumes a straight line relationship between the independent and dependent variables. However, in some cases the relationship may not be linear, in which case a linear model would not accurately reflect the true relationship. This can result in a poor fit and unreliable results. For example, if the discharge data shows a non-linear relationship with summer annual mean discharge, a linear model may not capture the true relationship and would not be a suitable choice for this analysis. It may be necessary to use a non-linear model or a different type of analysis that better captures the relationship between the independent and dependent variables.

## Way wordier than I would have put it, but basically correct - Ben